---
title: "Weather and Corn Regressions"
author: "Daniel Cleveland"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
library(ggplot2)
```


## Weather Data Analysis

### Load the PRISM daily maximum temperature
```{r tmax data, message = FALSE, warning = FALSE}
# daily max temperature
# dimensions of the data:  counties x days x years
prism <- readMat("data/prismiowa.mat")


#look at all temp for county #1 in first year of data (1981):
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]  #(shows that 1981 was not a leap year.)
plot(1:366, t_1981_c1, type = "l")

ggplot()+
  geom_line(mapping = aes(x=1:366, y=t_1981_c1))+
  theme_bw()+
  xlab("day of year") +
  ylab("daily maximum temperature (°C)")+  #use Alt+0176 as keyboard shortcut to get the degree symbol!!!
  ggtitle("Daily Maximum Temperature for County #1 in Iowa")

  
```


```{r tidying up, message = FALSE, warning = FALSE}
# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 
                                       1:366,
                                       prism$years)

# converting 3D matrxi to a dataframe:
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf) # tidying up

```


## Temperature Trends

### Summer Temperature Trends:  Winneshiek County

```{r temp trends, message = FALSE, warning = FALSE}
# converting some factors into dbl:
tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

# creating variable for containing summer temp values of county 191
winnesummer <- tmaxdf %>%
  filter(countyfp == 191 & doy >= 152 & doy <= 243) %>%
  group_by(year)%>%
  summarize(meantmax = mean(tmax))
```


```{r plotting summer temps, message = FALSE, warning = FALSE}
ggplot(winnesummer, mapping = aes(x = year, y = meantmax))+
  geom_point()+
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)")+
  geom_smooth(method = lm)

lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)


```

## Winter Temperatures trends:  Winneshek County

```{r getting winter temps for Winneshek, message = FALSE, warning = FALSE}
# creating variable for containing winter temp values of county 191
winnewinter <- tmaxdf %>%
  filter(countyfp == 191 & (doy <= 59 | doy >= 335) & !is.na(tmax)) %>%
  group_by(year)%>%
  summarize(meantmax = mean(tmax))
```


```{r plotting winter temps for Winneshek, message = FALSE, warning = FALSE}
ggplot(winnewinter, mapping = aes(x = year, y = meantmax))+
  geom_point()+
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)")+
  geom_smooth(method = lm)
```


```{r looking linear regression for winter, message = FALSE, warning = FALSE}
lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)

```


## Multiple Regression -- Quadratic Time Trend

```{r, message = FALSE, warning = FALSE}
winnewinter$yearsq <- winnewinter$year^2

lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
#summary(lm_winntertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

ggplot(winnewinter)+
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted))+
  theme_bw() +
  labs(x = "year", y = "tmax")


```


### Download NASS corn yield data

```{r corn yield download and convert to numeric, message = FALSE, warning = FALSE, results = 'hide'}

## set our API key with NASS
nassqs_auth(key = "6C0A99BB-3CB7-3E99-A699-B9DB7FDB8D16")

## parameters to query on
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN",
              prodn_practice_desc = "ALL PRODUCTION PRACTICES",
              year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

#convert to numeric
cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

```


```{r clean and filter dataset, message = FALSE, warning = FALSE}
#clean and filter dataset
cornyields <- select(cornyieldsall,county_ansi, 
                     county_name, yield, year)%>%
  filter(!is.na(county_ansi) & !is.na(yield))

# make a tibble
cornyields <- tibble(cornyields)

```


## Assignment

### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?

### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 

### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.

### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

### Bonus: Find a package to make a county map of Iowa displaying some sort of information about yields or weather. Interpret your map.

### Bonus #2: Challenge question - map trends in corn yields by county across Iowa. Interpret your map.

